A Lyapunov Analysis of Momentum Methods \\ in Optimization

### 2016-10-26 LyapDraft.tex
small corrections

%\begin{small}
\begin{subequations}\label{Eq:LyapAnal}
\begin{align}
 \frac{d}{dt}D_h\left(x, Z_t\right)  &= \frac{d}{dt}\left( h(x) - h(Z_t) - \langle \nabla h(Z_t), x- Z_t\rangle \right) \notag\\
& = - \langle \nabla h(Z_t), \dot Z_t\rangle -\left\langle \frac{d}{dt} \nabla h(Z_t),x - Z_t\right\rangle  + \langle \nabla h(Z_t), \dot Z_t\rangle \notag\\ 
 &= - \left\langle \frac{d}{dt} \nabla h\left(Z_t\right), x - Z_t\right\rangle \, \notag\\%\label{eq:apply-identity}\\
&=    \left\langle \frac{d}{dt}\left(e^{\beta_t}\right)\nabla f(X_t), x  - X_{t} - \frac{1}{\dot \beta_t} \dot X_{s} \right\rangle  \, \label{eq:apply-dynamics}   \\
& =  \frac{d}{dt}\left(e^{\beta_t}\right) \langle \nabla f(X_t), x - X_t \rangle\, - e^{ \beta_t} \langle \nabla f(X_t), \dot X_t\rangle dt\, \notag   \\
& = \frac{d}{dt} \left(e^{\beta_t}\right) \langle \nabla f(X_t), x - X_t \rangle\,  -  \frac{d}{dt} \left(e^{ \beta_t} f(X_t)\right)\,   + \frac{d}{dt} \left(e^{\beta_t}\right)  f(X_t)  \notag\\
&= \frac{d}{dt}\left( e^{\beta_t}\right) [f(X_t)+ \langle \nabla f(X_t), x - X_t \rangle] -\frac{d}{dt} \left(e^{ \beta_t} f(X_t)\right)\,  \notag  \\ %\label{Eq:Line9}\\
& \leq   \frac{d}{dt}\left( e^{\beta_t}\right) f(x) - \frac{d}{dt} \left(e^{ \beta_t} f(X_t)\right)\, \label{eq:apply-convexity}\\
& = - \frac{d}{dt} \left(e^{\beta_t}\left(f(X_t) - f(x)\right)\right) \label{Eq:cond1}.
\end{align}
\end{subequations}
%\end{small}
Here~\eqref{eq:apply-dynamics} uses the momentum dynamics~\eqref{Eq:ELH} and~\eqref{Eq:ELZ}. The inequality~\eqref{eq:apply-convexity} follows from the convexity of $f$. Rearranging terms and taking $x = x^\ast$, we have show that the function~\eqref{Eq:E} has nonpositive derivative for all $t$ and is hence a Lyapunov function for the dynamical system~\eqref{Eq:EL}.

@@

E_{k+1} - E_k & =  f(x_{k+1}) - f(x_k) - \mu\langle z_{k+1} - z_k, x^\ast - z_{k+1}\rangle - \frac{\mu}{2} \|z_{k+1} - z_k\|^2 \\

@@

E_{k+1} - E_k \leq - \tau_k \left( f(x_{k+1}) - f(x^\ast) + \frac{\mu}{2}\|x^\ast - z_{k+1}\|^2\right) = -\tau_k E_{k+1}.

@@

E_{k+1} - E_k &= f(y_{k+1}) - f(y_k) - \mu \langle z_{k+1} - z_k, x^\ast - z_{k+1}\rangle - \frac{\mu}{2}\|z_{k+1} - z_k\|^2\\

@@

E_{k+1} - E_k \leq -\tau_k E_k + \varepsilon_k 

@@

E_{k+1} - E_k &= f(y_{k+1}) - f(y_k) - \mu \langle z_{k+1} - z_k, x^\ast - z_{k+1}\rangle - \frac{\mu}{2}\|z_{k+1} - z_k\|^2\\

@@

E_{k+1} - E_k \leq -\tau_k E_k

>>>>

%\begin{small}
\begin{subequations}\label{Eq:LyapAnal}
\begin{align}
 \frac{d}{dt}D_h\left(x, Z_t\right)  &= \frac{d}{dt}\left( h(x) - h(Z_t) - \langle \nabla h(Z_t), x- Z_t\rangle \right) \notag\\
& = - \langle \nabla h(Z_t), \dot Z_t\rangle -\left\langle \frac{d}{dt} \nabla h(Z_t),x - Z_t\right\rangle  + \langle \nabla h(Z_t), \dot Z_t\rangle \notag\\ 
 &= - \left\langle \frac{d}{dt} \nabla h\left(Z_t\right), x - Z_t\right\rangle \, \notag\\%\label{eq:apply-identity}\\
&=    \left\langle \frac{d}{dt}\left(e^{\beta_t}\right)\nabla f(X_t), x  - X_{t} - \frac{1}{\dot \beta_t} \dot X_{s} \right\rangle  \, \label{eq:apply-dynamics}   \\
& =  \frac{d}{dt}\left(e^{\beta_t}\right) \langle \nabla f(X_t), x - X_t \rangle\, - e^{ \beta_t} \langle \nabla f(X_t), \dot X_t\rangle dt\, \notag   \\
& = \frac{d}{dt} \left(e^{\beta_t}\right) \langle \nabla f(X_t), x - X_t \rangle\,  -  \frac{d}{dt} \left(e^{ \beta_t} f(X_t)\right)\,   + \frac{d}{dt} \left(e^{\beta_t}\right)  f(X_t)  \notag\\
&= \frac{d}{dt}\left( e^{\beta_t}\right) [f(X_t)+ \langle \nabla f(X_t), x - X_t \rangle] -\frac{d}{dt} \left(e^{ \beta_t} f(X_t)\right)\,  \notag  \\ %\label{Eq:Line9}\\
& \leq   \frac{d}{dt}\left( e^{\beta_t}\right) f(x) - \frac{d}{dt} \left(e^{ \beta_t} f(X_t)\right)\, \label{eq:apply-convexity}\\
& = - \frac{d}{dt} \left(e^{\beta_t}\left(f(X_t) - f(x)\right)\right) \label{Eq:cond1}.
\end{align}
\end{subequations}
%\end{small}
Here~\eqref{eq:apply-dynamics} uses the momentum dynamics%~\eqref{Eq:EL}.
~\eqref{Eq:ELH} and~\eqref{Eq:ELZ}. 
The inequality~\eqref{eq:apply-convexity} follows from the convexity of $f$. Rearranging terms and taking $x = x^\ast$, we have show that the function~\eqref{Eq:E} has nonpositive derivative for all $t$ and is hence a Lyapunov function for the family of momentum dynamics~\eqref{Eq:EL}.

@@

\tilde E_{k+1} -\tilde  E_k & =  f(x_{k+1}) - f(x_k) - \mu\langle z_{k+1} - z_k, x^\ast - z_{k+1}\rangle - \frac{\mu}{2} \|z_{k+1} - z_k\|^2 \\

@@

\tilde E_{k+1} -\tilde  E_k \leq - \tau_k \left( f(x_{k+1}) - f(x^\ast) + \frac{\mu}{2}\|x^\ast - z_{k+1}\|^2\right) = -\tau_k \tilde E_{k+1}.

@@

\tilde E_{k+1} - \tilde E_k &= f(y_{k+1}) - f(y_k) - \mu \langle z_{k+1} - z_k, x^\ast - z_{k+1}\rangle - \frac{\mu}{2}\|z_{k+1} - z_k\|^2\\

@@

@@

\tilde E_{k+1} -\tilde  E_k &= f(y_{k+1}) - f(y_k) - \mu \langle z_{k+1} - z_k, x^\ast - z_{k+1}\rangle - \frac{\mu}{2}\|z_{k+1} - z_k\|^2\\

@@

\tilde E_{k+1} - \tilde E_k \leq -\tau_k\tilde  E_k
